---
title: "Calculus 2 Prediction"
author: "Jackie Budka"
date: "10/8/2021"
output: html_notebook
---

# Load the relevant packages

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(glue)
library(scales)
library(ggrepel)
library(DT)
library(readr)
library(janitor)
```

# Import the csv data and convert the names into snake case
```{r load csv}
df1a <- read_csv("CU_ACAD_DEPT_WHEN_OFF_ENRL_22740.csv", 
    skip = 1) %>% clean_names()
```
# Select useful columns and filter out irrelevant courses and course types
```{r all Calc I and II sections}
df1b <- df1a %>%
  select(course_id:catalog,title,term_7,term_8,tot_enrl,component) %>%
  filter(subject == "MA", title == "Calculus I" | title == "Calculus II", component == "LEC",grepl('Fall|Spring', term_8))%>%
  group_by(title,term_8)%>%
  summarise(enroll_per_term = sum(tot_enrl)) %>%
ungroup()
```
```{r isolate relevant Calc I and II sections (not needed)}
df1c <- df1b %>%
  filter((title == "Calculus I" & grepl('Fall',term_8)) | (title == "Calculus II" & grepl('Spring',term_8)))
```

# Separate the variable that contains the year and term and rename the class title
```{r Split and rename}
df1d <- df1b %>%
  separate(term_8, c("term", "year")) %>%
  mutate(
    title = case_when(
      title == "Calculus I" ~ "Calc 1",
      title == "Calculus II" ~ "Calc 2",
    ))
```

# Create a new variable called aca_year that is a range of years pertaining to an academic year
```{r academic year}
df1e <- df1d %>%
  mutate(year = as.integer(year)) %>%
  mutate(
    aca_year = case_when(
      term == "Fall" ~ str_c((year-2000),"-",(year-1999)),
      term == "Spring" ~ str_c((year-2001),"-",(year-2000))
    )
  )%>%
select(-c(year))

```


# Join the term and class section variables to create one variable called class_term
```{r Join term and section}
df1f <- df1e %>%
  mutate (
    class_term = 
      str_c((title)," ",(term))
    )%>%
select(-c(title,term))%>%
group_by(class_term)
```

# Pivot the data to use the values of enrollment per term versus the class_term
```{r}
df1g <- df1f %>%
  pivot_wider(names_from = class_term, values_from = enroll_per_term)%>%
  clean_names()
```

# Isolate the current calc 1 fall class to make prediction
```{r find current prediction}
df1hvert <- df1g %>%
  filter(aca_year == "21-22")
```


# Take out incomplete data
```{r filter out incomplete data}
df1h <- df1g %>%
  
filter(!aca_year=="21-22")
```

# Plot Fall Calculus I class vs Spring Calculus II class
```{r Plot Calc 1 Fall vs Calc 2 of the following Spring}
ggplot(df1h, aes(calc_1_fall, calc_2_spring)) + 
  geom_point() + geom_smooth(method = lm) + geom_vline(xintercept=457)

```

